Formalization of Abstract Rewriting

### 2025-02-11 Relations/Paper/Well-foundedness.tex
New structure for the paper

\documentclass{scrartcl}
% \input{packages}

\usepackage{booktabs}

\input{macros}

\title{Formalizing Abstract Rewriting}
% \title{Abstract Rewriting from a Constructive Viewpoint}
% \title{A Constructive View on Abstact Rewriting}
\author{Sam Arkle}
\date{\today}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\quad\ding{51}}%
\newcommand{\xmark}{\quad\ding{55}}%

\newcommand{\thra}{\twoheadrightarrow}
\newcommand{\thla}{\twoheadleftarrow}
\newcommand{\thlra}{{\thla\!\!\thra}}

\newcommand{\down}[1]{{#1}{\downarrow}}

\begin{document}

\renewcommand{\arraystretch}{1.3}

\maketitle
\begin{abstract}
We formalize in the Agda proof assistant the main results about Abstract Reduction Systems (ARS) given in the standard reference [Terese].
% Our contribution can be viewed as a development of abstract rewriting using constructive logic.
% , and is part of a larger project aiming to develop a library of programming language theory in Agda. This application leads us to give special focus to concepts related to termination and confluence.
We define a taxonomy of concepts related to termination and confluence and investigate their relationships between each other and their classical counterparts.
We identify, and eliminate where possible, the use of classical logic in the proofs of standard ARS results.
% For example, the study of strong normalization leads us to multiple inequivalent
% formalizations of the notion of a well-founded relations, with subtle distinctions.
Our analysis leads to refinements and mild generalizations of classical termination and confluence criteria.
\end{abstract}

\tableofcontents

\section{Introduction}
This is a report on the progress I have made during the summer of 2024 working with Dr. Polonsky \footnote{Assisstant Professor, Department of Computer Science, Appalachian State University.}. I have been writing on the topic of relations, working primarily from Term Rewriting Systems (TeReSe) \footnote{(Term Rewriting Systems. (2003). United Kingdom: Cambridge University Press.)}.
The objective of this work has been to both develop a formal understanding of relations and to build a library in Agda \footnote{\href{https://github.com/DrPolonsky/LAM/tree/main/Relations}{Github repository for this project's code}} that will be used as a foundation to further work in theoretical computer science.
\\
I have been focussed on Chapter 1 of TeReSe which covers abstract reduction systems (ARS) as well as the appendix (sections 1-3) which provides the necessary mathematical background for the rest of the book (e.g. relations, ordinals, inductive definitions and the Knaster-Tarski lemma). Where the book has made use of classical logic I have investigated the possibility of using constructive alternatives.

\section{Definitions} \label{section:Definitions}
The most interesting work we have carried out in investigating Chapter 1 of TeReSe has resulted from trying to provide constructive proofs for Newman's lemma, the Knaster-Tarski lemma, theorem 1.2.2 and theorem 1.2.3. Before exploring those proofs the following definitions are essential.

The following definitions are taken from [TeReSe], Chapter 1,
with minimal adaptations for consistency.

\begin{dfn}
  An abstract reduction system (\emph{ARS}) is a structure $\mathcal{A} = (A, R_\alpha)$ where $A$ is a set and $R_\alpha$ is a set of binary relations.
\end{dfn}
A relation $R$ from the set $R_\alpha$ is refered to as a reduction (or rewrite) relation. For example, if we specify $R$ as the relation of being 'less than' ($<$) then $R\, 4\, 2$. \textcolor{red}{Right?}
\apcomment{Yes, but move this to a(n eventual) list of examples.}

We use $R$ or $\to$ for a single-step reduction and $R^*$ or $\tto$ for a multi-step reduction.

\sacomment{Should we be using $R^+$ rather than $R^*$?}

\apcomment{Thus, when $(a,b) \in \to_{(\alpha)}$, we say that there is a
\emph{$(\alpha-)$reduction} from $a$ to $b$.
When $(a,b) \in \tto_\alpha$, we say that there is a multi-step $\alpha$-reduction
from $a$ to $b$.}

\begin{dfn}
  An element $a \in A$ is \emph{R-weakly confluent} ($WCR_R$) if all single step reductions from $a$ can always converge
  via multi-step reductions to some common reduct.
  A relation $R$ is \emph{weakly confluent} (AKA weakly Church-Rosser (WCR)) if every $a \in A$ is R-weakly confluent.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{R-confluent} ($CR_R$) if all multi-step reductions from $a$ can always converge
  via multi-step reductions to some common reduct.
  A relation $R$ is \emph{confluent} (AKA Church-Rosser (CR)) if every $a \in A$ is R-confluent.
\end{dfn}

\begin{figure}[h]
  \centering
\[  \begin{tikzpicture}[node distance=1.5cm, auto]
    % Nodes
    \node (top) {a};
    \node (left) [below left of=top] {b};
    \node (right) [below right of=top] {c};
    \node (bottom) [below right of=left] {d};

    % Arrows
    \draw[->] (top) -- (left);
    \draw[->] (top) -- (right);
    \draw[->>] (left) -- (bottom);
    \draw[->>] (right) -- (bottom);
\end{tikzpicture}
\hspace{3cm}
\begin{tikzpicture}[node distance=1.5cm, auto]
  % Nodes
  \node (top) {a};
  \node (left) [below left of=top] {b};
  \node (right) [below right of=top] {c};
  \node (bottom) [below right of=left] {d};

  % Arrows
  \draw[->>] (top) -- (left);
  \draw[->>] (top) -- (right);
  \draw[->>] (left) -- (bottom);
  \draw[->>] (right) -- (bottom);
\end{tikzpicture}
\]
\caption{On the left is an example of a weakly confluent relation. On the right is an example of a confluent relation.}
\end{figure}

\begin{dfn}
  \emph An element $a \in A$ is a {normal form} if there exists no reduction to any other element in $A$.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-weakly normalizing} ($WN_{R}$)if $a \tto b$ for some normal form $b \in A$.

  A relation $R$ is weakly normalizing (WN) if every $a \in A$ is $R$- weakly normalizing.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-strongly normalizing} ($SN_R$) if every reduction sequence starting from $a$ is finite.

  A relation $R$ is strongly normalizing (SN) if every
  $a \in A$ is $R$-strongly normalizing.
\end{dfn}

The above definition of SN is taken from [TeReSe]. When working in Agda, we make use of the alternative definition presented in [TeReSe] ($\leftarrow$ is accessibly well-founded).

\subsection{Closure operations on relations}

A relation $R$ has the following properties when the stated conditions are met:
\begin{dfn}
  \emph{reflexive} {- $\forall x \in A : x \to x $}
\end{dfn}
\begin{dfn}
  \emph{transitive} {- $\forall x, y, z \in A : ( x \to y \land y \to z ) \implies x \to z $}
\end{dfn}
\begin{dfn}
  \emph{symmetric} {- $\forall x, y \in A : x \to y \implies y \to x$}
\end{dfn}

\begin{dfn}
  The \emph{closure} of $R$ is the smallest relation that contains $\to$ and satisfies a given property.
\end{dfn}

\begin{dfn}
  The \emph{reflexive closure} of $R$ is the smallest relation $\to^r$ that contains $\to$ and is reflexive.
\end{dfn}
\begin{dfn}
  The \emph{symmetric closure} of $R$ is the smallest relation $R^s$ that contains $R$ and is symmetric.
\end{dfn}
\begin{dfn}
  The \emph{transitive closure} of $R$ is the smallest relation $R^+$ that contains $R$ and is transitive.
\end{dfn}
\begin{dfn}
  The \emph{reflexive transitive closure} of $R$ is the smallest relation $R^*$ that contains $R$ and is both reflexive and transitive.
\end{dfn}
\begin{dfn}
  The \emph{equivalence relation} of $R$ is the smallest relation $R^=$ that contains $R$ and is reflexive, transitive, and symmetric.
\end{dfn}

\begin{dfn}
  A relation $R$ has the \emph{weak normal form property} (WNFP) if
 $a \tto b$ and $a \tto c$  $\implies$
 $c \tto b$, for all $a,c \in A$ and all normal form $b \in A$.
\end{dfn}

\begin{dfn}
  A relation $R$ has the \emph{normal form property} (NFP) if
 $a \to^= b$  $\implies$
 $a \tto b$, for all $a \in A$ and any normal form $b \in A$.
\end{dfn}

Two elements $a, b \in A$ being equivalent is denoted $a \equiv b$.

\begin{dfn}
  A relation $R$ has the \emph{unique normal form property} (UN) if
  $a \to^= b$  $\implies$
  $a \equiv b$, for all $a \in A$ and any normal form $b \in A$.
\end{dfn}

\begin{dfn}
  A relation $R$ has the \emph{unique normal form property with respect to reduction} (UN$^ \to$) if
  $a \twoheadleftarrow  \cdot  \tto b$  $\implies$ $a \equiv b$, for all normal forms $a,b \in A$.
\end{dfn}

Let $A$ be a set.
\begin{dfn}
  A \emph{sequence (in $A$)} is a function from the natural numbers to $A$:
  \begin{align*}
    &s : \nat \to A \\
    &s = (s_0,s_1,s_2,\dots)
  \end{align*}
\end{dfn}

\begin{dfn}
  A sequence is \emph{$R$-increasing} if every term is $R$-related to its preceding term.
\end{dfn}

\begin{dfn}
  A sequence is \emph{bounded} if there exists an element $a$ to which all elements of the sequence reduce.
\end{dfn}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
      % Define the nodes inside the region f
      \node (f0) at (0, 0) {$f_0$};
      \node (f1) at (1, 0) {$f_1$};
      \node (dots) at (2, 0) {$...$};

      % Draw the arrows between the nodes inside the region f
      \draw[->] (f0) -- (f1);
      \draw[->] (f1) -- (dots);
      \draw[->] (f1) -- (dots);

      % Define the nodes outside the region f
      \node (a) [below=of dots, yshift=-.25cm] {\parbox{2cm}{\centering $a$ }};

          % Draw the arrows between the nodes inside and outside the region f sds

          \draw[->>] (f0) -- (a);
          \draw[->>] (f1) -- (a);
          \draw[->>] (dots) -- (a);
      \end{tikzpicture}
      \caption{An example of a bounded sequence.}

    \end{figure}

\begin{dfn}
  An element $a \in A$ has the \emph{recurrence property} if $a \tto b \implies b \tto a$ for all $b \in A$.
\end{dfn}
\sacomment{I've been using $\tto$ as interchangable with $R *$ (as in the above definition).
Is that okay or am I going to have to clarify why they are interchangable (are they not interchangable?)?}
\apcomment{It's okay, but later we will need to make sure we are using them consistently.  Say that in Agda code, $\to$ is denoted by $R$.}

\begin{dfn}
  A relation $R$ has the \emph{recurrence property} (RP) if every
  bounded, increasing sequence has a recurrent element.
\end{dfn}

    \begin{figure}[h]
      \centering
      \begin{tikzpicture}
          % Define the nodes inside the region f
          \node (f0) at (0, 0) {$f_0$};
          \node (f1) at (1, 0) {$f_1$};
          \node (dots) at (2, 0) {$...$};
          \node (fi) at (3, 0) {$f_i$};
          \node (fi+1) at (4, 0) {$f_{i+1}$};
          \node (dots2) at (5,0) {$...$};

          % Draw the arrows between the nodes inside the region f
          \draw[->] (f0) -- (f1);
          \draw[->] (f1) -- (dots);
          \draw[->] (dots) -- (fi);
          \draw[->] (fi) -- (fi+1);
          \draw[->] (fi+1) -- (dots2);

          % Define the nodes outside the region f
          \node (a) [below=of dots, yshift=-.25cm] {\parbox{1cm}{\centering $a$ }};

              % Draw the arrows between the nodes inside and outside the region f sds

              \draw[->>] (f0) to [bend right] (a);
              \draw[->>] (f1) to [bend right] (a);
              \draw[->>] (dots) -- (a);
              \draw[->>] (fi) to [bend left=10] (a);
              \draw[->>] (fi+1) to [bend left] (a);
              \draw[->>] (dots2) to [bend left] (a);
              \draw[->>] (a) to [bend left=10] (fi);

          \end{tikzpicture}
          \caption{An example of a relation with the recurrence property.}

        \end{figure}
\begin{dfn}
  A relation $R$ has the \emph{weak recurrence property} (RP-) if
  for every increasing sequence with bound $a$,
  there is an element $i$ of the sequence such that $a \tto f (i)$.
\end{dfn}

\begin{dfn}
  A relation $R$ has the \emph{recurrence property} (RP) if
 $a \tto b$ and $a \tto c$  $\implies$
 $c \tto b$, for all $a,c \in A$ and all $b \in A$ with the recurrence property.
\end{dfn}
\sacomment{This mirrors our Weak normal form property but for recurrence. Which would be a suitable name?}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-weakly recurrent} ($WR_{R}$)if $a \tto b$ for some recurrent element $b \in A$.

  A relation $R$ is weakly recurrent (WR) if every $a \in A$ is $R$- weakly recurrent.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-strongly recurrent} ($SR_R$) if either $a$ itself is recurrent or every element one $R$ step
  from $a$ is also strongly recurrent.

  A relation $R$ is strongly recurrent (SR) if every
  $a \in A$ is $R$-strongly recurrent.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-strongly recurrent} ($SR_R$) if every reduction sequence starting from $a$ reaches a recurrent element in a finite number of steps.

  A relation $R$ is strongly recurrent (SR) if every
  $a \in A$ is $R$-strongly recurrent.
\end{dfn}
\sacomment{We have the two contender definitions for SR above.}

\begin{dfn}
  \emph{Cofinality property}
\end{dfn}
\sacomment{Still not sure on how to define cofinality.}

\begin{dfn}
  A relation $R$ is \emph{finitely branching} (FB) if, for all $a \in A$, the set of one step reducts of $a$ is finite.
\end{dfn}

\subsection{Recurrence property and increasing relations}

In [TeReSe], a relation $R \sse A \times A$ is called \emph{increasing} if there
exists a ``size function'' $|-| : A \to \nat$ satisfying, for all $x, y \in A$,
 \[ Rxy \to |x| < |y| \]

\begin{prop}
  If $R$ is increasing, then no infinite $R$-increasing sequence is bounded.
\end{prop}
\begin{proof}
  Let $(a_i) = a_0 \to a_1 \to a_2 \to \cdots$ be an $R$-increasing sequence
  and suppose that $(a_i)$ is bounded.  Let $b$ be the bound, so that $a_i \to b$ for all $i$.

  If $m = |b|$ then we have
  \[ 0 \le |a_0| < |a_1| < |a_2| < \cdots < |a_m| < |b| = m \]
  % \[ m > |a_m| > |a_{m-1}| > \cdots > |a_0| \ge 0 \]
  which is a contradiction, since there are only $m-1$ integers strictly less than $m$.
\end{proof}

It follows that every relation that is increasing in the sense of [TeReSe]
vacuously satisfies the RP condition.

\begin{cor}
  If $R$ is increasing, then $R$ has the recurrence property.
\end{cor}

\section{Classical Principles}
Principles we are considering
\begin{itemize}
  \item Decidability of equality on $A$
  \item Decidability of the relation $R$
  \item When relevant, decidability of the given precicate $P$
  \item Not-not-closure of being accessible
  \item Not-not-closure of $R$
  \item Not-not-closure of $P$
  \item Decidability of whether a given element is an $R$-normal form
  \item When relevant, local/element-wise versions of the above
  \item Markov's principle:
  \begin{itemize}
    \item In $wfMin_0 \to wfSeq$.

    Given a sequence $s : \nat \to A$, let $E(a) = \exists n \in \nat. s n \equiv a$.

    Need: $\forall a. \lnot\lnot E(a) \to E(a)$, i.e.,
    \[\lnot\lnot \big(\exists n \in \nat. s n \equiv a\big)
          \to \big(\exists n \in \nat. s n \equiv a\big)\]
  \end{itemize}
  \item When an element is not SN, it transitions to one which is not SN either.
  (Use in the proof of Theorem 1.2.3)
\end{itemize}

\section{Implications}
{\small
We now look into the interrelation of the various properties defined in Section~\ref{section:Definitions}. Our aim is to expand on the
connections made in Figure 1.8 of [TeReSe] (p.19).

The following are the implications we have proven in Agda, along with a reference to the name of the proof in Agda. Where appropriate we indicate
which classical assumptions were used as part of the proof.

\begin{itemize}
  \item SN and WCR $\implies$ CR

\texttt{NewmansLemma : $\mathnormal{\forall}$ \{R : $\mathbb{R}$ A\} $\to$ SN R $\to$ weakly-confluent R $\to$ confluent R}

We have also provided a proof elementwise making use of the classical principal
about the decidability of $R$ being minimal.

\texttt{SN$\land$UN$\to$CRelem :\newline $\forall$ (R : $\mathbb{R}$ A) $\to$ decMin (~R R) $\to$ $\forall$ x $\to$ is R -SN x $\to$ is R -UN x$\to$ is R -CR x}

\sacomment{How to one line this consistently and in a pretty way?}

\item{CP $\implies$ CR}

\texttt{iv : CP R $\to$ confluent R}

This is theorem 1-2-3 iv in [TeReSe]

\item{WN and UN $\implies$ CR}

\texttt{ii : WN R $\times$ UN R $\to$ confluent R}

This is theorem 1-2-2 ii in [TeReSe]

\item{WN and UN $\implies$ bounded}

\texttt{i : WN R $\to$ UN R $\to$ bounded R}

\sacomment{Going to have to find a way to neatly represent the implications}

\sacomment{note to self, tidy up below with references to agda code}

\item{WN and UN$\to$ $\implies$ bounded}

\texttt{i+ : WN R $\to$ UN$\to$ R $\to$ bounded R}

\item{WN and UN and RP $\implies$ wfSeq - ($\sim$ R)}

\texttt{ii3- :  WN R $\to$ UN R $\to$ RP R $\to$ isWFseq- ($\sim$R R)}

\item {CR $\implies$ WCR}

\texttt{CR$\to$WCR : CR R $\to$ WCR R }

\item {NF $implies$ UN}
\sacomment{Need to prove}

\item {SN and WCR $\implies$ UN and WN}
\sacomment{Need to proove}
\texttt{}

\item {Ind and Inc $\implies$ SN} \sacomment{I don't think we have
a proof of this, should we include it, include it with a note that we're taking [TeReSe]'s word on it?}

\item {WCR, WN, and, Inc $\implies$ SN}

\item {WCR, WN, and RP $\implies$ SN}

\item {CR $\implies$ NF $\implies$ UN}

\texttt{i-4 : confluent R $\to$ NFP R $\to$ UN R}

\item {CR $^{\leq1} \implies$ CR}

\texttt{iv : CP R $\to$ confluent R}

\end{itemize}
}

\subsection{Double-check: What has been proven}
\begin{itemize}
  \item \sacomment{List all the constructive proofs we've put together!}
  \item SN implies bounded
  \item RP- implies RP
  \item The infinite haircomb refutes $WN \land RP \to \omega$-bounded.  But it is not WCR.
  \item We have formalized proposition 1.1.9, 1.1.10, 1.1.11 from [TeReSe].
  \item We have a proof for Theorem 1.2.1 (Newlman's Lemma) (though we have just one proof of multiple possible proofs).
  \item We have a proof for all parts of Theorem 1.2.2
  \item We have a proof for Theorem 1.2.3 i, iv.
  \item Reduction closure properties for SN, NF, UN, recurrent property.

  So, does $WN, RP, WCR$ imply $\omega$-bounded?
  Answer: yes, that is what we have proved classically.
\end{itemize}

\subsection{Hierarchy}
The properties of strong normality and confluence are the two properties of greatest interest. As such, there is
value in determining which other properties combine to imply one these properties. This motivated the development of the
following hierarchies.

\subsubsection{Confluence hierarchy}

\begin{center}

  \begin{tikzpicture}[auto,
    arrowstyle/.style={->, line width=1pt, >={Latex[length=3mm, width=2mm]}, shorten >=2pt}]
    % A box style
    \tikzstyle{boxnode} = [draw, rectangle, minimum width=3cm, minimum height=1cm, text centered]

    % Place the nodes vertically (top to bottom)
    \node (Confluent) [boxnode] at (0,0) {Confluent};
    \node (RP)        [boxnode, below=2cm of Confluent] {Recurrence Property};
    \node (WN)        [boxnode, below=2cm of RP] {Weak Normal Form Property};
    \node (UN)        [boxnode, below=2cm of WN] {Unique Normal Form Property};

    % Draw arrows downwards
    \draw[arrowstyle] (Confluent) -- (RP);
    \draw[arrowstyle] (RP) -- (WN);
    \draw[arrowstyle] (WN) -- (UN);
  \end{tikzpicture}
\end{center}

\subsubsection{Normalization hierarchy}

\begin{tikzpicture}[node distance=2cm, auto,
  % Define a thicker arrow with larger arrowhead and some shortening
  arrowstyle/.style = {->, line width=1pt, >={Latex[length=3mm, width=2mm]}, shorten >=2pt}
]
% Define a box style
\tikzstyle{boxnode} = [draw, rectangle, minimum width=3cm, minimum height=1cm, text centered]

% Define the nodes
\node (SN) [boxnode] at (0,0) {Strongly Normal};
\node (WN) [boxnode, below left=2cm and 2cm of SN] {Weakly Normal};
\node (SR) [boxnode, below right=2cm and 2cm of SN] {Strongly Recurrent};
\node (WR) [boxnode, below of=SN, yshift=-4cm] {Weakly Recurrent};

% Draw the arrows to form a diamond
\draw [arrowstyle] (SN) -- (WN);
\draw [arrowstyle] (SN) -- (SR);
\draw [arrowstyle] (WN) -- (WR);
\draw [arrowstyle] (SR) -- (WR);
\end{tikzpicture}

The property of being strongly recurrent implies the property of being weakly recurrent only when there is also the classical property
of decidability of minimality for a given element.

In the following table red text indicates that classical properties are required for the implication to hold. Blue indicates we are currently unsure.

\begin{table}[h!]
  \centering
  \caption{Local implications}
  \begin{tabular}{|l|c|c|c|c|c|}
  \hline
  \textbf{CR} & CR, CE-9  & CR, CE-6  & CR , CE-9  & CR, SN  & CR , SN \\
  \hline
  \textbf{RP} & CR, CE-9  & CR, CE-6  & \red{CR} , CE-9 & CR, SN   & \red{CR} , SN \\
  \hline
  \textbf{WNFP} & CE-10, CE-9  & CR, CE-6  & CE-11 , CE-9 & CR, SN   & \red{CR} , SN \\
  \hline
  \textbf{UN} & CE-3, CE-12  & CE-3, CE-6  & CE-3, CE-9  & CE-3, CE-3  & \red{CR} , SN \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Local implications - WCR }
  \begin{tabular}{|l|c|c|c|c|c|}
  \hline
  \textbf{CR} & CR, CE-9  & CR, {CE-6}  & CR , {CE-9}  & CR, SN  & CR , SN \\
  \hline
  \textbf{RP} & CR, CE-9  & CR, {CE-6}  & \red{CR} , {CE-9} & CR, SN   & \red{CR} , SN \\
  \hline
  \textbf{WNFP} & CE-10.1, CE-9  & CR, {CE-6}  & {CE-11} , {CE-9} & CR, SN  & \red{CR} , SN \\
  \hline
  \textbf{UN} & CE-3.1, CE-12  & {CE-3.1}, {CE-6}  & {CE-3.1}, {CE-9}  & {CE-3.1}, {CE-3.1}  & \red{CR} , SN \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Global implications}
  \begin{tabular}{|l|c|c|c|c|c|}
  \hline
  \textbf{CR} & CR, {CE-9}  & CR, {CE-6}  & CR , {CE-9}  & CR, SN  & CR , SN \\
  \hline
  \textbf{RP} & CR, {CE-9}  & CR, {CE-6}  & \red{CR} , {CE-9} & CR, SN   & \red{CR} , SN \\
  \hline
  \textbf{WNFP} & {CE-11}, {CE-9}  & CR, {CE-6}  & CE-11 , {CE-9} & CR, SN   & \red{CR} , SN \\
  \hline
  \textbf{UN} & {CE-3}, {CE-12} and CE-9  & CR, {CE-6}  & CE-3, {CE-9}  & CR, SN  & \red{CR} , SN \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Global implications - WCR}
  \begin{tabular}{|l|c|c|c|c|c|}
  \hline
  \textbf{CR} & CR, {CE-9}  & CR, {CE-6}  & CR , {CE-9}  & CR, SN  & CR , SN \\
  \hline
  \textbf{RP} & CR, {CE-9}  & CR, {CE-6}  & {CR} , {CE-9} & CR, SN   & CR , SN \\
  \hline
  \textbf{WNFP} & CE-1.2, {CE-9}  & CR, {CE-6}  & {CR} , {CE-9} & CR, SN   & CR , SN \\
  \hline
  \textbf{UN} & CE-1.1, {CE-9}  & CR, {CE-6}  & {CR}, {CE-9}  & CR, SN  & CR , SN \\
  \hline
  & \textbf{WR} & \textbf{WN} & \textbf{SR} & \textbf{WN}$\boldsymbol{\land}$\textbf{SR} & \textbf{SN} \\
  \hline
  \end{tabular}
\end{table}

\newpage
\section{Counterexamples}
The following counterexamples are used to rule out certain implications between properties from holding.
\subsection*{Counterexample 1}
\begin{center}

  \begin{tikzcd}[node distance=2cm, auto]
      % Nodes
      \node (a) {a};
      \node (b) [right of=a, xshift=1cm] {b};
      \node (c) [right of=b, xshift=1cm] {c};
      \node (d) [right of=c, xshift=1cm] {d};

      % Arrows
      \draw[->] (b) to [bend left] (c);
      \draw[->] (c) to [bend left] (b);
      \draw[->] (b) -- (a);
      \draw[->] (c) -- (d);
  \end{tikzcd}

\end{center}
Counterexample to:
\begin{itemize}
  \item WCR $\implies$ CR
  \item WCR  $\implies$ WNFP
  \item \mbox{WCR R $\to$ $\forall$ a b $\to$ is R -NF b $\to$ (R *) a b $\to$ $\forall$ c $\to$ R a c $\to$ (R *) c b }
  \item WCR and WN $\implies$ CR
  \item WCR and WN $\implies$ SN
  \item IOW and WCR $\implies$ WNFP \sacomment{We need to define IOW in the report}
\end{itemize}

\subsection*{Counterexample 1.1}
\begin{center}

  \begin{tikzcd}[node distance=2cm, auto]
      % Nodes
      \node (a) {a};
      \node (b) [right of=a, xshift=1cm] {b};
      \node (c) [right of=b, xshift=1cm] {c};
      \node (d) [right of=c, xshift=1cm] {d};

      % Arrows
      \draw[->] (b) to [bend left] (c);
      \draw[->] (c) to [bend left] (b);
      \draw[->] (b) -- (a);
      \draw[->] (c) -- (d);
      \draw[->, loop] (d) to (d);

  \end{tikzcd}

\end{center}
Counterexample to:
\begin{itemize}
  \item (Global) WCR, UN, and WR $\implies$ WNFP
\end{itemize}

\subsection*{Counterexample 1.2}
\begin{center}

  \begin{tikzcd}[node distance=2cm, auto]
      % Nodes
      \node (a) {a};
      \node (b) [right of=a, xshift=1cm] {b};
      \node (c) [right of=b, xshift=1cm] {c};
      \node (d) [right of=c, xshift=1cm] {d};

      % Arrows
      \draw[->] (b) to [bend left] (c);
      \draw[->] (c) to [bend left] (b);
      \draw[->] (b) -- (a);
      \draw[->] (c) -- (d);
      \draw[->, loop] (d) to (d);
      \draw[->, loop] (a) to (a);

  \end{tikzcd}

\end{center}
Counterexample to:
\begin{itemize}
  \item (Global) WCR, WNFP, and WR $\implies$ RP
\end{itemize}

\subsection*{Counterexample 2}
\begin{center}
  \begin{tikzcd}
      % Nodes
      \node (x) {a};
      \node (n) [left of=x, xshift=-1cm] {n \in NF};
      \node (y) [right of=x, xshift=1cm] {b};
      \node (z) [right of=y] {c};

      % Arrows
      \draw[->] (x) -- (n);
      \draw[->] (x) -- (y);
      \draw[->] (y) to [bend left] (z);
      \draw[->] (z) to [bend left] (y);
  \end{tikzcd}
\end{center}

Counterexample to:
\begin{itemize}
  \item WCR and an increasing sequence where the first elelement of the sequence reduces to an element in normal form $\to$
  the sequence is bounded.
\end{itemize}

\subsection*{Counterexample 3}
\begin{center}
  \begin{tikzcd}
      % Nodes
      \node (x) {a};
      \node (n) [left of=x, xshift=-1cm] {n \in NF};
      \node (y) [right of=x, xshift=1cm] {b};
      \node (z) [right of=y] {c};

      % Arrows
      \draw[->>] (x) -- (n);
      \draw[->>] (x) -- (y);
      \draw[->] (y) to [bend left] (z);
      \draw[->] (z) to [bend left] (y);
  \end{tikzcd}
\end{center}

Counterexample to:
\begin{itemize}
  \item WN$\downarrow\subseteq$WN (WN is not closed under reduction) \sacomment{Is this the right way to express that elementwise WN doesn't imply global WN?}
  \item WN$\downarrow$UN$\rightarrow\subseteq$WN :

  UN$\rightarrow\,$R $\to \forall$ x $\to$ is R-WN x $\to$ $\forall$ y $\to$ (R *) x y $\to$ is R-WN y

  \item $\forall$ (R : $\mathscr{R}$ A) $\to$ $\forall$ is R -WN x $\to$ is R -UN x $\to$ is R -CR x
  \item $WN_R$ and $UN_R$ $\implies$ $CR_R$
  \item (Local) WR and UN $\implies$ NP
  \item (Local) WN and UN $\implies$ SR

\end{itemize}

\subsection*{Counterexample 3.1}
\begin{center}
  \begin{tikzcd}
      % Nodes
      \node (x) {a};
      \node (a') [above of=x] {a'};
      \node (n) [left of=x, xshift=-1cm] {n \in NF};
      \node (y) [right of=x, xshift=1cm] {b};
      \node (z) [right of=y] {c};

      % Arrows
      \draw[->>] (x) -- (n);
      \draw[->] (a') -- (x);
      \draw[->] (a') -- (n);
      \draw[->>] (x) -- (y);
      \draw[->] (y) to [bend left] (z);
      \draw[->] (z) to [bend left] (y);
  \end{tikzcd}
\end{center}

Counterexample to:
\begin{itemize}
  \item (Local) WCR, WR, and UN $\implies$ NP
  \item (Local) WCR, WN and UN $\implies$ SR

\end{itemize}

\subsection*{Counterexample 4}

\begin{center}
  \begin{tikzcd}
      % Nodes
      \node (x) {a};

      % Arrows
      \draw[->, loop] (x) to (x);
  \end{tikzcd}
\end{center}
Counterexample to:
\begin{itemize}
  \item bounded and RP $\implies$ SN
  \item bounded and RP $\implies$ WN
\end{itemize}

\subsection*{Counterexample 5}
An infinite sequence
  \begin{center}
      \begin{tikzpicture}
      % Define the nodes
      \node (a) at (0, 0) {$f_0$};
      \node (b) at (1, 0) {$f_1$};
      \node (dots) at (2, 0) {$...$};

      % Draw the arrows
      \draw[->] (a) -- (b);
      \draw[->] (b) -- (dots);
      \end{tikzpicture}
    \end{center}
Counterexample to:
\begin{itemize}
  \item WN
  \item bounded
  \item RP $\implies$ bounded
  \item WCR $\implies$ WN
  \item CR $\implies$ WN
  \item CR and CP $\implies$ WN
  \item CR and CP $\implies$ bounded
\end{itemize}

\subsection*{Counterexample 6}
An infinite sequence with a bound that is a normal form.
\begin{center}

  \begin{tikzpicture}
      % Define the nodes inside the region f
      \node (f0) at (0, 0) {$f_0$};
      \node (f1) at (1, 0) {$f_1$};
      \node (dots) at (2, 0) {$...$};

      % Draw the arrows between the nodes inside the region f
      \draw[->] (f0) -- (f1);
      \draw[->] (f1) -- (dots);
      \draw[->] (f1) -- (dots);

      % Define the nodes outside the region f
      \node (a) [below=of dots, yshift=-.25cm] {\parbox{2cm}{\centering $n\in NF$ }};

          % Draw the arrows between the nodes inside and outside the region f sds

          \draw[->] (f0) -- (a);
          \draw[->] (f1) -- (a);
          \draw[->] (dots) -- (a);
      \end{tikzpicture}

    \end{center}
Counterexample to:
\begin{itemize}
  \item WN $\implies$ SN
  \item WN and CR $\implies$ SN
  \item WN, CR, and bounded $\implies$ SN
  \item WN, UN, and bounded $implies$ SN
  \item (Local) WN and UN $\implies$ SR
  \item (Local) WN and NP $\implies$ SR
  \item (Local) WN and RP $\implies$ SR
  \item (Local) WN and CR $\implies$ SR
  \item (Local) WN, SR, and NP $\implies$ SN
  \item (Local) WN, SR, and RP $\implies$ SN
  \item (Local) WN, SR, and CR $\implies$ SN
\end{itemize}

\subsection*{Counterexample 7}
An infinite sequence with a bound that is a normal form where each element
of the sequence has a single step relation to the initial element of the sequence.
\begin{center}
  \begin{tikzpicture}
      % Define the nodes inside the region f
      \node (f0) at (0, 0) {$f_0$};
      \node (f1) at (1, 0) {$f_1$};
      \node (dots) at (2, 0) {$...$};

      % Draw the arrows between the nodes inside the region f
      \draw[->] (f0) -- (f1);
      \draw[->] (f1) -- (dots);
      \draw[->] (f1) -- (dots);
      \draw[->] (f1) to [bend right=20] (f0);
      \draw[->] (dots) to [bend right=50] (f0);

      % Define the nodes outside the region f
      \node (a) [below=of dots, yshift=-.25cm] {\parbox{2cm}{\centering $n\in NF$ }};

          % Draw the arrows between the nodes inside and outside the region f sds

          \draw[->>] (f0) -- (a);
          \draw[->>] (f1) -- (a);
          \draw[->>] (dots) -- (a);
      \end{tikzpicture}

    \end{center}
Counterexample to:
\begin{itemize}
  \item RP
\end{itemize}

\subsection*{Counterexample 8}
An "infinite haircomb"
\begin{center}
  \begin{tikzpicture}
      % Define the nodes inside the region f
      \node (f0) at (0, 0) {$f_0$};
      \node (f1) at (1, 0) {$f_1$};
      \node (f2) at (2, 0) {$f_2$};
      \node (dots) at (3, 0) {$...$};

      \node (n0) at (0, -1) {$n_0$};
      \node (n1) at (1, -1) {$n_1$};
      \node (n2) at (2, -1) {$n_2$};
      \node (dots2) at (3, -1) {$...$};

>>>>

\section{Well-foundedness}
\label{sec:Well-foundedness}
